\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage [autostyle, english = american]{csquotes}

\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{conj}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{conditions}{Conditions}[section]

%\newcommand{\defv}[1]{\textbf{\textit{#1}}}



\title{The local optimality of the double lattice packing}
\author{Yoav Kallus and W\"oden Kusner}
%\date{}							% Activate to display a given date or no date

\begin{document}
\bibliographystyle{plain}
%\maketitle
%\abstract
%This paper shows that the dense double lattice construction of Kuperberg and Kuperberg is locally optimal for polygons in the full space of packings.


\section{The construction of the densest double lattice packing}

\subsection{Double lattices}

\begin{definition}
    A chord of a convex body $K$ is a line segment whose endpoints lie on the boundary of $K$.
    A chord is an affine diameter if there is no longer chord parallel to it.
\end{definition}

\begin{definition}
    An inscribed parallelogram is a half-length parallelogram in the direction $\theta$ if
    one pair of edges is parallel with the line through the origin at an angle $\theta$ above
    the $x$-axis and their length is half the length of an affine diameter parallel to them.
\end{definition}

Note that any two half-length parallelograms in the direction $\theta$ have equal
area, and we can define that area as a function $A(\theta)$ of the direction.

\begin{definition}
    A cocompact discrete subgroup of the Euclidean group consisting of translations
    and point reflections is a \text{double lattice} if it includes at least one
    point reflection.
\end{definition}

A double lattice is generated by a lattice and a point reflection, or alternatively
by three point reflections.

\begin{theorem}[Kuperberg and Kuperberg]\label{thmkk}%strict convexity is not necessary (see Mount)
    For a convex $K$, an admissible double lattice of smallest mean area has
    mean area $4\min_\theta A(\theta)$ and is generated by reflection about
    the vertices of a half-length parallelogram.
\end{theorem}

%By a sequence of approximations, this gives that for convex bodies $K$ that are not strictly convex, there
%exists a double lattice packing of maximal density that generated a minimum area extensive parallelogram in $K$.

Kuperberg and Kuperberg make use of extensive parallelograms, inscribed parallelograms with edge length greater than half the affine diameter in their edge directions, but make use of half-length parallelograms. Mount gives an explicit proof that, when minimizing the volume of an extensive parallelogram, it suffices to consider only the half-length parallelograms associated with affine diameters as a function of $\theta.$

From the definition of an extensive parallelogram, one observes that this is exactly the condition that produces an admissible, tight double lattice packing, and that the density of the packing increases as the area of the extensive parallelogram decreases.  Thus, there is a direct correspondence between the admissible motions of the double lattice packing and the motions of inscribed, extensive parallelograms. 

 For example, sliding motions a double lattice packing that preserves density correspond to motions of an inscribed extensive parallelogram that preserve area.

%include images.

For our purposes, it is illustrative to consider the algorithm used find the half-length parallelogram of minimal area not only as a method of finding the best double lattice packing, but also as a way to explore part of the configuration space of double lattices. More specifically, the behavior of the affine diameter is fairly well behaved but non-trivial as $\theta$ increases while the behavior of the vertices of the half parallelogram are described by an {\it interspersing property}; they are non-decreasing functions $[0,2\pi)\rightarrow [0,2\pi).$  %they are a non-decreasing function, they never back track....

We begin with a simple proposition:
\begin{proposition}
there is an affine diameter of a convex polygon K in every direction that meets a vertex.
\end{proposition}

This is a consequence of the convexity K, for if a chord does not meet a vertex, it either lies within an edge or it meets two edges.  If it lies within an edge, it is not an affine diameter.  If it meets two edges, it is possible increase its length while parallel translating it in a non-decreasing direction of the cone defined by the two edges until it meets a vertex. 
%picture

From this, it is possible to determine an initial affine diameter in a particular direction.  Furthermore, this is done by geometrically satisfying procedure of extending a set of parallel rays in a from all vertices into the interior of K and selecting the longest.  However, it does not give 

\subsection{Walking around vertices}




%critical angles



%\begin{remark}
%Furthermore,  IMVT should say that ever point is met by an affine diameter as well, so every vertex is met by one.? Combining these satisfies the intuition that one may 
%\end{remark}



%continuous sweeping vs finding the minimum area intermediate configuration...
%can introduce the notion of stability here

%Having found the optimal configuration, it is clear that the condition of being a strict minimal configuration and the  




The densest double lattice packing of a convex polygon $K$ can be constructed in time proportional to the number of vertices by an algorithm of Mount \cite{Mount1991}. The goal of this paper is to show that this configuration is not only a local maximum
of density among double lattices, but is in fact a local maximum in a broader sense, strong extremality.










\bibliography{ngon}
\end{document}  